---
title: "Assignment 3"
author: "Kaz Sakamoto"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, error = FALSE)
```

In this assignment we will be scraping data from Wikipedia. Please submit a rscript (.r file) with your submission. Title should be Lastname_Firstname_assignment3.r and submit to canvas.

Use these urls for scraping.
```{r}

## let's create two variables url1 and ulr2 with our wikipedia pages
url1 <- "https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population"
url2 <- "https://en.wikipedia.org/wiki/List_of_countries_by_GDP_(PPP)_per_capita"

```


## Task 1

for `url1`, name the table `pop` and rename columns to PopRank, Country, Population, Date, PopPercent, Source

## Task 2

for `url2` name the table `gdp` and rename columns to gdpRank, Country, GDP

# Task 3

take all the "-" in the `PopRank` and `gdpRank` columns in the tables and fill down with the value 
from the row above. Use the function `fill()` from the tidyr package.
https://tidyr.tidyverse.org/reference/fill.html

# Task 4

clean the country names and join gdp and pop together using `full_join()` from the dplyr package. Name the new table joinedDF. https://dplyr.tidyverse.org/reference/join.html


# Task 5

what are the top 10 countries by GDP in order? (greatest to smallest)

# Task 6

what are the top 10 countries by population in order? (greatest to smallest)