---
title: "Assignment2"
author: "Kaz Sakamoto"
date: "3/19/2018"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, error = FALSE)
```

In this assignment we will be scraping data from Wikipedia. Please submit a rscript (.r file) with your submission. Title should be Lastname_Firstname_assignment2.r and submit to canvas.

You can run the code snippet from below to help you get going!

```{r}
# thi will download htmltab if you don't have it
if (!require(htmltab)) install.packages('htmltab')

# let's load the library
library(htmltab)

## let's create two variables url1 and ulr2 with our wikipedia pages
url1 <- "https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population"
url2 <- "https://en.wikipedia.org/wiki/List_of_countries_by_GDP_(PPP)_per_capita"

# take the second table in population page
pop <- htmltab(url1,2)

# tale the second table in GDP page
gdp <- htmltab(url2,2)

```


## Task 1

rename pop columns to PopRank, Country, Population, Date, PopPercent, Source

## Task 2

rename gdp columns to gdpRank, Country, GDP

# Task 3

take all the "-" in the Rank columns in both tables and fill down with the value 
from the row above

# Task 4
clean the country names and join gdp and pop together

# Task 5
what are the top 10 countries by GDP in order? (greatest to smallest)

# Task 6
what are the top 10 countries by population in order? (greatest to smallest)