---
title: "Assignment 7"
author: "Kaz Sakamoto"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, error = FALSE)
```

In this assignment we will be continue working with Rmarkdown and R. Please submit a html document as your submission. Title should be Lastname_Firstname_assignment7.html and submit to canvas.

# Task 1

This week well be using the `Boston` data set you can check up the data in this
[file](https://cran.r-project.org/web/packages/MASS/MASS.pdf) for the column meanings.

```{r}
library(MASS)
data(package="MASS")
df <- Boston
```

# Task 2

Make some exploratory graphs and clean the data if need be.

# Task 3

create a `tree` with the Boston data predicting `medv` with all other variables. What
do you see as an result?

# Task 4

create a train and test split with you data.
try a bunch of `mtry`s with this data set on `randomForest` still predicting for `medv` with all other variables. 
Compare the train results with the test results (MSE). What does the visual output look like? What number 
of mtrys is best?

# Task 5

now train the `randomForest` on your entire data with the `mtry` you believe is best. What do the results look like?