---
title: "Assignment 8"
author: "Kaz Sakamoto"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, error = FALSE)
```

In this assignment we will be continue working with Rmarkdown and R. Please submit a html document as your submission. Title should be Lastname_Firstname_assignment8.html and submit to canvas.

# Task 1

This week we will revist rvest and Wikipedia. I want you to scrape the text in the paragraph tags in a few different pages.
Look for "id-bodyContent", then "p" and grab the text. 
```{r}
urbanPlanning <- "https://en.wikipedia.org/wiki/Urban_planning"
urbanDesign <- "https://en.wikipedia.org/wiki/Urban_design"
urbanEconomics <- "https://en.wikipedia.org/wiki/Urban_economics"
Architecture <- "https://en.wikipedia.org/wiki/Architecture"
```

# Task 2

combine the text into one tibble with columns,`topic`(should include the page topic), `text`.

# Task 3

Tokenize the text and remove stop words from the text. I want you to visualize the top 20 words from each
page. explain what you see.

# Task 4

Now I would like you to plot the word frequencies against urban planning. I'll also want to see the correlations
between the pages and explanation of what you see. 

# Task 5

Using the udpipe library I would like you to compare keywords from each article using
`as_phrasemachine` and `keywords_phrases`. Explain what you see.