---
title: "Intro to R"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Basic Math
R is great for statistics but can handle basic math operations. Let's have a try:

```{r calculator}
1 + 1
1 + 2 + 3
3 * 7 * 2
4 / 2
4 / 3
```

## PEMDAS 

Let's see how order affects how code is run:
```{r pemdas}
4 * 6 + 5
(4 * 6) + 5
4 * (6 + 5)
```

# Variables

## Assigning variables
When assignining variables you can use that "<-" or "=" to assign variables:
```{r variable-assignment}
x <- 2
x
y = 5
y

3 -> z
z

a <- b <- 7
a
b

assign("j", 4)
j
```

## Removing 

Sometimes you need to remove a variable from your workspace:
```{r removing-vars}
j
rm(j)
# now it is gone
j
```

## Case Sensitive 

In R variables are case sensitive:
```{r case-sensitive-vars}
theVariable <- 17
theVariable
THEVARIABLE
```

# Data Types 

## Numeric

As we saw earlier R can handle numeric data types:
```{r numeric}
x <- 2
x
class(x)
is.numeric(x)
```

## Integer

Integers are also numeric data types but do not have any decimal places:
```{r integer}
i <- 5L
i
is.integer(i)
is.numeric(i)
```

## Integer Division

R will convert integers to numeric when needed, for example when you are mixing
types in multiplcation or division:
```{r int-division}
class(4L)
class(2.8)
4L * 2.8
class(4L * 2.8)

class(5L)
class(2L)
5L / 2L
class(5L / 2L)
```

## Character Data 

Another data type is character which can take in strings:
```{r characters}
x <- "data"
x
```

let's take a count of the number of characters in our variable:
```{r nchar-intro}
nchar(x)
nchar("hello")
nchar(3)
nchar(452)
```


## Dates

Dates are also important and distinct, there are Date (Days) and POSIXct (seconds)
from Janurary 1, 1970:
```{r create-dates}
date1 <- as.Date("2012-06-28")
date1
class(date1)
as.numeric(date1)

date2 <- as.POSIXct("2012-06-28 17:42")
date2
class(date2)
as.numeric(date2)
```


## Logical

Booleans are another type which stores True or False. They are stored numerically
as 1 = True, 0 = False:
```{r true-false-mult}
TRUE * 5
FALSE * 5
```

```{r logical-check}
k <- TRUE
class(k)
is.logical(k)
```

We can also compare variables as logicals:
```{r comparison-example}
# does 2 equal 3?
2 == 3
# does 2 not equal three?
2 != 3
# is two less than three?
2 < 3
# is two less than or equal tothree?
2 <= 3
# is two greater than three?
2 > 3
# is two greater than or equal to three?
2 >= 3
# is "data" equal to "stats"?
"data" == "stats"
# is "data" less than "stats"?
"data" < "stats"
```

# Vectors 

## Vector Creation 

A vector is a collection of the same data type:
```{r create-first-vector} 
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
x
```

## Vector Operations

What if we want to multiply each element in our vector by 3:
```{r vector-elem-mult}
x * 3
```

As you saw earlier, vectorized operations are great:
```{r vector-examples}
x + 2
x - 3
x / 4
x^2
sqrt(x)
```

## Vector Range Creation
Using a ":" will also to create a range between the two numbers:
```{r consecutive-vector}
1:10
10:1
-2:3
5:-7
```

## Vector Operations 

We can even run oeprations against two vectors of equal length:
```{r vector-two-operations}
# create two vectors of equal length
x <- 1:10
y <- -5:4
# add them
x + y
# subtract them
x - y
# multiply them
x * y
# divide them--notice division by 0 results in Inf
x / y
# raise one to the power of the other
x^y
# check the length of each
length(x)
length(y)
# the length of them added together should be the same
length(x + y)
```
Unlike the example above, if one vector is shorter it will get recycled. 
If the longer vector is not a multiple of the shorter a warning will be thrown:
```{r vector-recycling}
x + c(1, 2)
x + c(1, 2, 3)
```
Comparissions also works as vectors:
```{r logical-vectors}
x <= 5
x > y
x < y
```


## Character Vectors 

We've been looking at numeric vectors, but what about character vectors?
```{r nchar-vector,tidy=FALSE}
q <- c("Hockey", "Football", "Baseball", "Curling", "Rugby",
       "Lacrosse", "Basketball", "Tennis", "Cricket", "Soccer")
nchar(q)
nchar(y)
```

## Vector Elements

by using the "[" we can select elements by their order number:
```{r vector-elements}
q[1]
q[1:2]
q[c(1, 4)]
```

## Vector Names 

You can provide names for elements in your vector:
```{r vector-names}
# provide a name for each element of an array using a name-value pair
c(One="a", Two="y", Last="r")

# create a vector
w <- 1:3
# name the elements
names(w) <- c("a", "b", "c")
w
```

# Calling Functions

## Mean and Sum 

let's take the mean and the sum of our previous vector:
```{r mean-simple}
mean(x)
sum(x)
```

## Getting help 
Using a "?" will help you get to the help page for a function:
```{r binary-help,eval=FALSE,tidy=FALSE}
?mean
```

Not sure what the name of the function is? try apropos:
```{r apropos}
apropos("mea")
```

# Missing Data {#MissingData page=RfE50}

## NA

You will most likely have to deal with NAs some point in your work:
```{r na-example}
z <- c(1, 2, NA, 8, 3, NA, 3)
z
is.na(z)
```

calcluating functions with NAs might not work how you expect:
```{r na-mean}
mean(z)
mean(z, na.rm = TRUE)
```


Here is an example with character data:
```{r na-types}
zChar <- c("Hockey", NA, "Lacrosse")
zChar
is.na(zChar)
```

## NULL

Null differs from NA which is represented as another element in the vector. Null
is more similar to nothingness:
```{r null-disappear}
z <- c(1, NULL, 3)
z
```

```{r null-test}
d <- NULL
is.null(d)
is.null(7)
```
