---
title: "Data Munging"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Reading in Data

Let's download subway.csv and subway.xlsx from canvas and add it to a folder called Data in your project directory.

### Reading in CSV

CSVs are probably one of the most common data formats that you will work with.
To read in data we will use the "readr" package.

```{r read-subway-do}

## here we are going to find the root_file 
root <- rprojroot::find_rstudio_root_file()
## we are going to set the data file path here
dataDir <- file.path(root, 'Data')

# load readr
library(readr)
# read in csv
subway <- read_csv(file=file.path(dataDir, "subway.csv"))
```

How about reading files from online? just add the url in `read_csv()`

```{r read-subway,eval=FALSE}
subway_online <- read_csv(file="http://ml-course.kazsakamoto.com/Data/subway.csv")
```

Let's take a look at what subway loos like

```{r subway-head-show}
head(subway)
```

## Read Excel

Sometimes you will have to work with excel files. There is a special package for that.

```{r load-readxl,echo=FALSE}
## Load in readxl
library(readxl)
```

let's now use our `dataDir` again and load the excel file and see the available sheets.

```{r see-sheets-show, eval=FALSE, warning = FALSE, error=FALSE}
excel_sheets(file.path(dataDir, 'subway.xlsx'))
```

we can load the actual table with the specified sheet for now we'll just index it with the first sheet.
```{r read-subway-with-readxl-do,eval=FALSE}
subwayXL <- read_excel(file.path(dataDir, 'subway.xlsx'), sheet=1)
```


## Pipes

let's load the library magrittr.

```{r load-magrittr}
library(magrittr)
```
  
pipes are very useful and easier to read. Let's compare viewing the head of the
data with and without pipes.

Often it's the case where functions will be nested together and in those circumstances you'll have to read from inside-out. Pipes allows us to read from left to right!
```{r head}
head(subway)
subway %>% head
```

what if we wanted to get only the top four observations? let's add n = 4 as an argument.

```{r}
head(subway, n=4)
subway %>% head(n=4)
```


What if we wanted to get just the fourth row? isn't pipes much easier to read?

```{r head-tail}
tail(head(subway, n=4), n=1)
subway %>% head(n=4) %>% tail(n=1)
```


## dplyr

"dplyr is a grammar of data manipulation, providing a consistent set of verbs that help you solve the most common data manipulation challenges"

here is the [cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf) if you want a reference guide.

### select

If we need to select ceration columns from our table there are a few ways to do it
in dplyr we use select
```{r dplyr-select}
# with brackets
subway[, c('Lat', 'Long')]
# with pipes
subway %>% select(Lat, Long)
# with pipes and column index
subway %>% select(5, 6)
# everything but lat and long
subway %>% select(-c(5, 6))
```

### filter

if we want to apply certain filters to our data, we can use filter to add
logical arguments
```{r dplyr-filter}
subway[subway$LINE == 'SIR', ]
subway %>% filter(LINE == 'SIR')
# and
subway %>% filter(LINE == 'SIR', Long < -74.18)
# and
subway %>% filter(LINE == 'SIR' & Long < -74.18)
# or
subway %>% filter(LINE == 'SIR' | Long < -74.18)
# or
subway %>% filter(LINE %in% c('SIR', '6'))
```

let's filter out nas from our names

```{r}
subway %>% filter(!is.na(NAME)) -> subway
```

### slice

If we want to select rows by index numbers, we can use slice
```{r dplyr-slice}
subway %>% slice(1:5)
subway %>% slice(1:10)
```


### mutate 

If you want to create a new column use the mutate function

```{r dplyr-mutate}
# pass / stairs
subway %>% mutate(pass / stairs)
subway %>% mutate(ratio = pass / stairs)

# writing back to original data.frame
subway %<>% mutate(ratio = pass / stairs)
subway
```

We can even include more than one column to be made
```{r dplyr-mutate-seq}
subway %>% mutate(passLog = log(pass), TwiceSize = ratio * 2)
```

## mutate_at and mutate_if
```{r}
subway %>% select(Lat,Long) %>% 
    mutate_if(is.numeric, funs(as.character))

subway %>% mutate_at(vars(starts_with("La")), funs(as.character))
```


### summarize

what if we want to make some summary stats?
```{r dplyr-sumamrize}
# average pass
subway %>% summarize(mean(pass))
# naming summary stats
subway %>% summarize(AvgPass = mean(pass))
```

### group_by 
like pivot tables, maybe you want to group your data?
```{r dplyr-group-by}
subway %>% group_by(borough) %>% 
    summarize(AvgPass=mean(pass))

subway %>% group_by(borough, stairs) %>% 
    summarize(AvgPass=mean(pass))
```

how about multiple summaries?
```{r dplyr-group-by-mult}
subway %>% 
    group_by(borough) %>% 
    summarize(AvgPass=mean(pass), AvgStairs=mean(stairs), TotalPass=sum(pass))

subwayWide <- subway %>% 
    group_by(borough, stairs) %>% 
    summarize(AvgPass=mean(pass), Count = n(), TotalPass=sum(pass)) 

subwayWide
```

### gather
```{r}
subwayLong <- subwayWide %>% 
    gather(key = variable, value = value, -borough, - stairs)

subwayLong
```


### spread
```{r}
subwayLong %>% spread(key = variable, value = value)
```


### unite
```{r}
subwayCoord <- subway %>% unite(Coord, Lat, Long, sep = " ")
```

### separate
```{r}
subwayCoord %>% separate(Coord, c("Lat", "Long"), sep = " ")
```

### arrange

```{r}
subway %>% arrange(pass)
subway %>% arrange(desc(pass))
## multiple ordering
subway %>% arrange(desc(pass), desc(stairs))
```

## Strings continued

concatenating strings can be an important task.

```{r build-paste-expression,tidy=FALSE}
message <- "Hello Bob, it is going to be 42 degrees today and sunny"

person <- "claire"
temp <- 78
condition <- "rainy"

paste("Hello ", person, ", it is going to be", temp, 
      "today and", condition)

paste("Hello ", person, ", it is going to be", temp, 
      "today and", condition, sep = "")
```

here is another way to do it with sprintf
```{r sprintf-sentence}
sprintf("Hello %s, it is going to be %s degrees today and %s", person, temp, condition)
```

what about over a vector?
```{r sprintf-sentence-vector}
sprintf("Hello %s, it is going to be %s degrees today and %s", c("Claire", "Bob", "Nick"), temp, condition)
```

## String Manipulation

```{r }
library(stringr)
# split the string
crossStreets <- str_split(string=subway$NAME, pattern=" & ")
head(crossStreets)
# combine them into one table
crossStreetsTable <- data.frame(Reduce(rbind, crossStreets))
head(crossStreetsTable)
# give the columns good names
names(crossStreetsTable) <- c("Street_1", "Street_2")
head(crossStreetsTable)
# bind the new columns onto the data.frame
cbind(subway, crossStreetsTable) %>% head
```

```{r }
# get the first 3 characters
str_sub(string=subway$NAME, start=1, end=5)
# get the 4rd through 8th characters
str_sub(string=subway$NAME, start=4, end=8)
```

## regular expressions

if you need to learn more about regular expressions take a look at this [site](https://regexone.com/)

remember in R you need another '\' to escape special characters
```{r }
Avenues <- str_detect(string=subway$NAME, pattern="Ave")
sum(Avenues, na.rm = TRUE)
Streets <- str_detect(string = subway$NAME, pattern = "St")
sum(Streets, na.rm = TRUE)
```

```{r }
# looks for r or R
subway %>% filter(str_detect(NAME, pattern="[r,R]"))
# begins with r or R
subway %>% filter(str_detect(NAME, pattern="^[r,R]"))
# ends with d
subway %>% filter(str_detect(borough, pattern="[d]$"))

## all punctuations which include !"#$%&’()*+,-./:;<=>?@[]^_`{|}~
subway %>% filter(str_detect(NAME, pattern = "[[::punct::]]"))
## periods
subway %>% filter(str_detect(NAME, pattern = "\\."))
## anything that has a space then capital E then a space or a period
subway %>% filter(str_detect(NAME, pattern = "\\s[E][\\s|\\.]"))
## anything that has a space then a alphabetic character then a space or period
subway %>% filter(str_detect(NAME, pattern = "\\s[A-z][\\s|\\.]"))

```

```{r}
Search <- str_detect(subway$LINE, "L")
sum(Search)
```

```{r}
# this looks for all numbers in subway line
str_extract_all(string=subway$LINE, "[0-9]")
```


```{r}
# replace the first digit seen with "x"
# \\d is the same as [0-9]
str_replace(string=subway$LINE, pattern="\\d", replacement="x") %>% head
str_replace_all(string=subway$LINE, pattern="\\d", replacement="x") %>% head
```


## rvest
To scrape html tables from websites, the package rvest is here to help.
```{r}
library(rvest)
```

Let's first read in a table. The World Bank has a table of one of its indicators [size of the economy](http://wdi.worldbank.org/table/WV.1) which we can scrape. We can use this link and read the html in
```{r}
wbURL <- read_html("http://wdi.worldbank.org/table/WV.1")
```
Next if we open up the inspector on chrome it'll open the html for the page. If you select the table we can copy the xpath which is '//*[@id="scrollTable"]'.
```{r}
econSize <- wbURL %>% 
    html_nodes(xpath = '//*[@id="scrollTable"]') %>% 
    html_table() %>% "[["(1)
```

We also have to get the column names.
```{r}
wbColNames1 <- wbURL %>% 
    html_nodes(".level0") %>% 
    html_nodes("th") %>% 
    html_text() %>% stringr::str_trim()

wbColNames1ext <- c(wbColNames1[-8], wbColNames1[7], rep(wbColNames1[8],2))

wbColNames2 <- wbURL %>% 
    html_nodes(".level1") %>% 
    html_nodes("th") %>% 
    html_text() %>% stringr::str_trim()

wbColNames3 <- wbURL %>% 
    html_nodes(".level2") %>% 
    html_nodes("th") %>% 
    html_text() %>% stringr::str_trim()

wbColNames <- paste(wbColNames1ext, wbColNames2, wbColNames3) %>% stringr::str_trim()
wbColNames[1] <- "Country"

colnames(econSize) <- wbColNames
```

Now you have the table!

```{r}
View(econSize)
```

## Google Maps Places API

Next we'll actually use an [API](https://developers.google.com/places/web-service/search) to gain data from Google Maps. This is an official way that companies/entities want to expose their data to others. Every user will need to register their key and use it when they access the API, you can get yours [here](https://developers.google.com/places/web-service/get-api-key). The API in this case is accessed through a url which is what shows up when you are actually using Google Maps!

First thing is to creat the api call, or in this case it's a url. The required parameters for a Place search is:

* key - your API key
* location - THe lat/long around your search
* radius - search distance in meters (max is 50,000 m)

optional parameters

* type - search restriction ([types supported](https://developers.google.com/places/web-service/supported_types))
* keyword - search word to be matched by Google content

If we are looking for Asisn restaurants around a 1,000 meter radius from Columbia
https://maps.googleapis.com/maps/api/place/nearbysearch/json?location=40.8075355,-73.9713274&radius=1000&type=restaurant&keyword=asian&key=YOUR_API_KEY


Let's think about making this algorithmic!
```{r}
library(RCurl)
lat <- 40.8075355
long <- -73.9713274
radius <- 1000
type <- "restaurant"
keyword <- "asian"
key <- "**YOUR_KEY**"

nearbySearch <- function(lat, long, radius, type, keyword, key){
    # this creates the api call
    url <- sprintf("https://maps.googleapis.com/maps/api/place/nearbysearch/json?location=%s,%s&radius=%s&type=%s&keyword=%s&key=%s",
            lat, long, radius, type, keyword, key)
    ## This reads in the url
    output <- getURL(url)
    ## Read in json and return the results object from the list
    outDF <- jsonlite::fromJSON(output)$results
    ## return outDF
    outDF
}

asianRest <- nearbySearch(lat, long, radius, type, keyword, key)
```
