---
title: "Rmarkdown and Plotting"
author: ""
date: ""
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# ggplot2

Now we will need to learn how to make some charts. Often times it is good for data scientists to understand
their data, as well as showing others their work, and it facilitates information gain.

First we will call the library and load in data.
```{r}
library(ggplot2)
download.file("http://kazsakamoto.com/data/NYC_Lots.csv", 
              "Data/NYC_Lots.csv")

root <- rprojroot::find_rstudio_root_file()
dataDir <- file.path(root, 'Data')

# we are now going to load one table in using read_csv
NYC_Lots <- readr::read_csv(file.path(dataDir, "NYC_Lots.csv"))

NYC_Lots %>% 
group_by(Borough) %>% 
sample_frac(.05) %>% filter(LotArea<20000) -> NYC_Lots
```

```{r}
# let's add comps to our function ggplot
ggplot(NYC_Lots)

# now we will add PctChng as our x value
ggplot(NYC_Lots, aes(x = LotArea))

# to create a histogram we need to add geom_histogram()
ggplot(NYC_Lots, aes(x = LotArea)) + geom_histogram()

# to create a density plot we use geom_density()
ggplot(NYC_Lots, aes(x = LotArea)) + geom_density()

# to change the color we will add this argument color='red' in geom_density
ggplot(NYC_Lots, aes(x = LotArea))+ geom_density(color = "red")

# that's great let's add fill='blue' to geom_density too
ggplot(NYC_Lots, aes(x = LotArea)) + geom_density(color='red', fill = "blue")

```

```{r}
## now we need to variables to plot
## we will add x=LotArea, y=TotalValue 

ggplot(NYC_Lots, aes(LotArea, TotalValue)) + geom_point()

## we can also add x=LotArea, y=TotalValue  in geom_point()
ggplot(NYC_Lots) + aes(LotArea, TotalValue)


# we will now color our scatterplot by color=Borough
ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue)) + geom_point(aes(color = Borough))


# here is an example of adding a trend line
ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue)) + 
geom_point(aes(color=Borough)) + geom_smooth(aes(color=Borough)) + 
geom_smooth(color='green')

# let's change the shape of our points add shape=1 to geom_point()
ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue, color=Borough)) + 
geom_point(shape = 1)

# we can also change the shape by a variable add shape=Borough now
ggplot(NYC_Lots, aes(x=LotFront, y=TotalValue, color=Borough)) + 
geom_point(aes(shape = Borough))

## Sometimes there might be too much data and lots of overlaping data
## we can change the transparency with alpha=1/4

ggplot(NYC_Lots, aes(x=LotFront, y=TotalValue, color=Borough)) + 
geom_point(shape=1, size=1, aplha = 1/4)

# to create a violin plot wi will add geom_violin()
ggplot(NYC_Lots, aes(x=Borough, y=TotalValue)) + geom_violin()

# we will now run through adding the points back in with geom_point()
ggplot(NYC_Lots, aes(x=Borough, y=TotalValue)) + geom_violin() + geom_point()

# adding geom_jitter() will help you visualize it better
ggplot(NYC_Lots, aes(x=Borough, y=TotalValue)) + geom_violin() + geom_jitter()

# again order matters so we will first place the jitter than the violin layer
ggplot(NYC_Lots, aes(x=Borough, y=TotalValue)) + geom_jitter() + geom_violin()

# and let's now add more layers to finish off our plot
ggplot(NYC_Lots, aes(x=Borough, y=TotalValue)) +
    geom_jitter(shape=1, size=1, alpha=1/3, width=.35, aes(color=LotArea)) + 
    geom_violin(alpha=1/2, color='grey50', draw_quantiles=c(.25, .5, .75)) +
    scale_color_gradient('Area', low='#56B1F7', high='#C91010') +
    labs(x='Borough', y='Value')


## we will now load ggthemes
library(ggthemes)

# let's assign our plot to a variable p
p <- ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue)) + geom_point(aes(color=Borough))

p + theme_minimal

# this time let's also add scale_color_economist()
p + theme_economist()

p + theme_economist() + scale_color_economist()


# let's look at theme_fivethirtyeight()
p + theme_fivethirtyeight()

p + theme_fivethirtyeight() + scale_color_fivethirtyeight()

# this time with theme_tufte()
p + theme_tufte() + scale_color_fivethirtyeight()

# theme_excel()
p + theme_excel()

# now with right colors using scale_color_excel()
p + theme_excel() + scale_color_excel()


## we can also break our plots by variables
## let's add facet_wrap(~ Borough) to create
## individual plots for each Borough
ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue, color=Borough)) + 
    geom_point(shape=1, size=1, alpha=1/4) + facet_wrap(~ Borough)

## So far we only used one variable Borough
## we can also plot against another variable
## let's add NumBldgs

ggplot(NYC_Lots, aes(x=LotArea, y=TotalValue, color=Borough)) + 
    geom_point(shape=1, size=1, alpha=1/4) +
    facet_grid(Borough ~ NumBldgs)
```

# Rmarkdown

# YAML Header {#yamlHeader page=RfES427}

YAML means Yet Another Markup Language

```YAML
---
title: "RMarkdown Primer"
author: "Your Name"
date: "Jan 12, 2019"
output: html_document
---
```

# Sections {#sections page=RfES428}

Sections are denoted with hashtags

```md
# Level 1 Heading

Regular text goes here. Blank lines create new paragraphs.

# Another Level 1 Heading

More text

## Level 2 Heading

This is a subsection under the level 1 heading

# Back to Level 1

This starts a new section
```

# Text Formatting {#text-formatting page=RfES429}

- Underscores for italic
- Two underscores for bold
- Three underscores for italic and bold

```md
_italic_

__bold__

___italic and bold___
```

_italic_

__bold__

___italic and bold___

# Lists {#lists page=RfES428}

## Unordered Lists {#lists-unordered page=RfES429}

A leading dash (`-`) creates a list element

List fruit

```md
- apples
- bananas
- oranges
- kiwi
```

- apples
- bananas
- oranges
- kiwi

## Ordered Lists {#lists-ordered page=RfES430}

### Numbered Lists {#lists-numbered page=RfES430}

A leading `1.` creates list elements

```md
1. New York
1. New Jersey
1. Connecticut
```

1. New York
1. New Jersey
1. Connecticut

### Lettered Lists {#lists-lettered}

A leading 'a.' creates a lettered list

```md
a. Coal
a. Wood
a. Gas
```

a. Coal
a. Wood
a. Gas

## Nested List {#lists-nested page=RfES430}

```md
1. NY
    a. Coal
    a. Wood
    a. Gas
        - Bakers Pride
        - Bari
    a. Electric
1. NJ
1. Philly
1. Detroit
```

1. NY
    a. Coal
    a. Wood
    a. Gas
        - Baker's Pride
        - Bari
    a. Electric
1. NJ
1. Philly
1. Detroit

# Links {#links page=RfES430}

```md
[My Website](www.ml-course.kazsakamoto.com)
```

[My Website](www.ml-course.kazsakamoto.com)

# Math {#math page=RfES430}

Markdown uses LaTeX math notation.

```md
$$
    \bar{x} = \sum_{i=1}^n \frac{x_i}{n}
$$
```

$$
    \bar{x} = \sum_{i=1}^n \frac{x_i}{n}
$$

```md
Inline equations look like $y = mx + b$.
```

Inline equations look like $y = mx + b$.

# Footnotes {#footnotes}

```md
We are learning markdown^[Markdown is a simple way of writing documents] on our way to learning RMarkdown.
```

We are learning markdown^[Markdown is a simple way of writing documents] on our way to learning RMarkdown.

# More YAML Options {#MoreYamlOptions page=RfES428}

Go to the top at add new options

```YAML
---
title: "RMarkdown Primer"
author: "Your Name"
date: "Jan 12, 2019"
output: 
    html_document:
        number_sections: yes
        toc: yes
        toc_float:
            collapsed: false
            smooth_scroll: true
        keep_md: false
        code_download: true
---
```

# Adding Custom sytling

```html
<style>
@import url('https://fonts.googleapis.com/css?family=Noto+Sans|VT323');

body{
  font-family: 'Noto Sans', sans-serif;
  font-size: 16px;
  line-height: 24px;
}

h1,h2,h3,h4 {
  font-family: 'VT323', monospace;
  background-color: #D4DAEC;
}

.container { width: 1000px; }


h2 {
  text-indent: 50px; 
}

g-table-intro h4 {
  text-indent: 0px;
}
</style>
```
